Binary logistic regression
=========================

TODO
-------------------------

 - link to associationOrder for `ROCR`, regressionOrdinal, regressionMultinom, regressionDiag for outliers, collinearity, crossvalidation

Install required packages
-------------------------

[`rms`](http://cran.r-project.org/package=rms)

```{r}
wants <- c("rms")
has   <- wants %in% rownames(installed.packages())
if(any(!has)) install.packages(wants[!has])
```

Descriptive model fit
-------------------------

### Simulate data
    
```{r}
set.seed(1.234)
N      <- 100
X1     <- rnorm(N, 175, 7)
X2     <- rnorm(N,  30, 8)
Y      <- 0.5*X1 - 0.3*X2 + 10 + rnorm(N, 0, 6)
Yfac   <- cut(Y, breaks=c(-Inf, median(Y), Inf), labels=c("lo", "hi"))
dfRegr <- data.frame(X1, X2, Yfac)
```

```{r rerRegressionLogistic01}
cdplot(Yfac ~ X1, data=dfRegr)
cdplot(Yfac ~ X2, data=dfRegr)
```

### Fit the model

```{r}
(glmFit <- glm(Yfac ~ X1 + X2,
               family=binomial(link="logit"), data=dfRegr))
```

```{r}
exp(coef(glmFit))
exp(confint(glmFit))
```

### Fit the model based on a matrix of counts

```{r}
total  <- sample(40:60, N, replace=TRUE)      ## unequal n_i
hits   <- rbinom(N, size=total, prob=0.4)
hitMat <- cbind(hits, total-hits)
glm(hitMat ~ X1 + X2, family=binomial(link="logit"))
```

### Fit the model based on a matrix with relative frequencies

```{r}
hitMatRel <- sweep(hitMat, 1, total, "/")
glm(hitMatRel ~ X1 + X2, weights=total, family=binomial(link="logit"))
```

### Fitted logits and probabilities

```{r rerRegressionLogistic02}
logitHat <- predict(glmFit, type="link")
plot(logitHat, pch=c(1, 16)[unclass(dfRegr$Yfac)])
abline(h=0)
```

```{r}
Phat <- fitted(glmFit)
Phat <- predict(glmFit, type="response")
head(Phat)
mean(Phat)
prop.table(table(dfRegr$Yfac))
```

Assess model fit
-------------------------

### Classification table

```{r}
thresh <- 0.5
Yhat   <- cut(Phat, breaks=c(-Inf, thresh, Inf), labels=c("lo", "hi"))
cTab   <- table(Yfac, Yhat)
addmargins(cTab)
sum(diag(cTab)) / sum(cTab)
```

### log-Likelihood, AUC, Somers's $D_{xy}$, Nagelkerke's pseudo $R^{2}$

```{r}
logLik(glmFit)
```

```{r}
library(rms)
lrm(Yfac ~ X1 + X2, data=dfRegr)
```

For plotting the ROC-curve, see `ROCR` in associationOrder

### McFadden and Cox & Snell seudo $R^{2}$

McFadden pseudo-$R^2$

```{r}
glmFit0 <- update(glmFit, . ~ 1)
LLf <- logLik(glmFit)
LL0 <- logLik(glmFit0)
as.vector(1 - (LLf / LL0))
```

Cox & Snell

```{r}
as.vector(1 - exp((2/N) * (LL0 - LLf)))
```

Nagelkerke

```{r}
as.vector((1 - exp((2/N) * (LL0 - LLf))) / (1 - exp(LL0)^(2/N)))
```

### Crossvalidation

`cv.glm()` function from package `boot`, see crossvalidation

Coefficient and model tests
-------------------------

### Individual coefficient tests

```{r}
summary(glmFit)
```

```{r results='hide'}
lrm(Yfac ~ X1 + X2, data=dfRegr)
```


### Model comparisons

```{r}
anova(glmFit0, glmFit, test="Chisq")
```

```{r results='hide'}
lrm(Yfac ~ X1 + X2, data=dfRegr)
```

```{r}
drop1(glmFit, test="Chi")
```

Detach (automatically) loaded packages (if possible)
-------------------------

```{r}
try(detach(package:rms))
try(detach(package:survival))
try(detach(package:splines))
```
